---
title: "The PTMods package: a package to handle post-translational modifications"
output:
    BiocStyle::html_document:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Description and usage of Spectra object}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
    %\VignettePackage{PTMods}
    %\VignetteDepends{PTMods}
---

```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

**Package**: `r Biocpkg("PTMods")`<br />
**Authors**: `r packageDescription("PTMods")[["Author"]] `<br />
**Last modified:** `r file.info("PTMods.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

# The unimod database

> Unimod is a public domain database, distributed under a
> [copyleft licence](http://www.gnu.org/licenses/license-list.html): "a
> copyright notice that permits unrestricted redistribution and
> modification, provided that all copies and derivatives retain the same
> permissions."
>
> The aim is to create a community supported, comprehensive database of
> protein modifications for mass spectrometry applications. That is,
> accurate and verifiable values, derived from elemental compositions,
> for the mass differences introduced by all types of natural and
> artificial modifications. Other important information includes any
> mass change, (neutral loss), that occurs during MS/MS analysis, and
> site specificity, (which residues are susceptible to modification and
> any constraints on the position of the modification within the protein
> or peptide).

Source: http://www.unimod.org/unimod_help.html


# Different types of PTM annotations


# Session information

```{r si}
sessionInfo()
```
