---
title: "The PTMods package: a package to handle post-translational modifications"
output:
    BiocStyle::html_document:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{The PTMods package}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
    %\VignettePackage{PTMods}
    %\VignetteDepends{PTMods}
---

```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

**Package**: `r Biocpkg("PTMods")`<br />
**Authors**: `r packageDescription("PTMods")[["Author"]] `<br />
**Last modified:** `r file.info("PTMods.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

Post-translational modifications (PTMs) are covalent modifications
that happen after protein biosynthesis and that can have a major
influence on protein function. PTMs are biologically very important
and can be identified in a high-throughput way using
mass spectrometry.

The `PTMods` package focuses on handling such PTMs and works in
conjuction with the `r BiocStyle::Biocpkg("PSMatch")` and
`r BiocStyle::Biocpkg("Spectra")` packages. It distributes PTMs from
the Unimod database and allows to convert PTM annotations between
multiple annotation formats.

```{r lib}
library(PTMods)
```

## Installation

The package can be installed from Bioconductor with

```{r install, eval = FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("PTMods")
```

Alternatively, altough this isn't recommended in general, it can also
be installed from Github.

```{r install2o, eval = FALSE}
BiocManager::install("RforMassSpectrometry/PTMods")
```

# The Unimod database

The `PTMods` package distributes the data from the Unimod database:

> Unimod is a public domain database, distributed under a
> [copyleft licence](http://www.gnu.org/licenses/license-list.html): "a
> copyright notice that permits unrestricted redistribution and
> modification, provided that all copies and derivatives retain the same
> permissions."
>
> The aim is to create a community supported, comprehensive database of
> protein modifications for mass spectrometry applications. That is,
> accurate and verifiable values, derived from elemental compositions,
> for the mass differences introduced by all types of natural and
> artificial modifications. Other important information includes any
> mass change, (neutral loss), that occurs during MS/MS analysis, and
> site specificity, (which residues are susceptible to modification and
> any constraints on the position of the modification within the protein
> or peptide).

Source: http://www.unimod.org/unimod_help.html

The package provides the following objects.

- The `modifications` dataframe describing `r
  nrow(PTMods::modifications)` PTMs from the unimod database:

```{r modifications}
head(modifications)
```

- The `aminoacids` dataframe describing `r nrow(PTMods::aminoacids)`
  amino acids from the unimod database:


```{r aminoacids}
head(aminoacids)
```
- The `elements` data.frame describing `r nrow(PTMods::elements)`
  chemical elements from the unimod database:

```{r elements}
head(elements)
```

Note that the goal of `PTMods` is not to place this information in a
biological context. There are other sources of biological information
about post-translational modifications:


- [RESID](https://proteininformationresource.org/resid/resid.shtml)
  for post-translational modifications
- [Uniprot/Swiss-Prot](https://www.uniprot.org/) protein sequence
  database
- [Prosite](https://prosite.expasy.org/) database of protein families
  and domains
- [Glycan Database](https://www.functionalglycomics.org/) from the
  Consortium for Functional Glycomics
- [PhosphoDB](https://phosphodb.hecklab.com/site/index) and
  [PhosphoSitePlus](https://www.phosphosite.org/homeAction.action) for
  phosphoryations
- ...

Other compilations of modifications with a focus on mass spectrometry:

- [Delta Mass](https://abrf.org/resources/delta-mass/)
- [FindMod](https://web.expasy.org/findmod/)
- ...

# Different types of PTM annotations

PTMs can be annotated using different syntaxes:

- The *deltaMass* syntax that describes the mass shift of an amino
  acid such as for example `"M[+15.994915]PEPTIDE"` for an oxidation.
- The *unimodId* that uses the unimod identifier such as for example
  `"M[UNIMOD:35]PEPTIDE"` for an oxidation.
- The modification's name, such as for example `"M[Oxidation]PEPTIDE"`.


The `convertAnnotion()` function can be used to convert between these
different syntaxes:

```{r convertAnnotion}
convertAnnotation("M[Oxidation]PEPTIDE", convertToStyle = "deltaMass")
convertAnnotation("M[Oxidation]PEPTIDE", convertToStyle = "unimodId")
convertAnnotation("M[+15.995]PEPTIDE", convertToStyle = "name")
```

# Session information

```{r si}
sessionInfo()
```